Long term cloud enhancement events of global solar irradiance over Thessaloniki, Greece

### 2024-06-12 article/article.tex
Commit 2024-06-12 10:14

The distribution of the number of CE events as a function of OIR is
shown in Figure\nobreakspace{}\ref{fig:ovir-distribution}. Apparently,
there is an inverse relation between the frequency of CE events and OIR
with an exponential-like decline. This is expected, as the stronger the
CE events become, the rarer are the conditions favoring the occurrence
of CE events. For the majority (over \(60\%\)) of the CE events the OIR
is below the long term average of \(39.9\,\text{W}/\text{m}^2\), while
about \(10\%\) of the events correspond to OIR larger than
\(100\,\text{W}/\text{m}^2\) and up to highest value of 300
W/m\textsuperscript{2}. This distribution is indicative of the magnitude
and the probability of the expected CE events over Thessaloniki. Similar
distribution of CE events, occurrences have been reported by
\citet{Vamvakas2020}, where the magnitude of OIR were higher due to the
location of the city of Patras, \(2.5^\circ\) closer to the equator.

@@

\caption{Distribution of CE over irradiance magnitude.}\label{fig:ovir-distribution}
\end{figure}

\hypertarget{groups-of-cloud-enhancement}{%
\subsection{Groups of cloud
enhancement}\label{groups-of-cloud-enhancement}}

In order to study the total duration of the CE events, we grouped the
one minute CE events to continuous CE groups (CEG). Thus, a CEG consists
of one or more successive CE cases, and can represent cloud enhancement
conditions spanning a duration of multiple minutes. We have identified
28468 groups of CE in the whole period of study, where the group with
the longest duration lasted 140 minutes on 07~July 2013. By examining
the frequency distribution of the CEG durations
(Figure\nobreakspace{}\ref{fig:ceg-duration-distribution}), we can
conclude, that although we detected some CEG with durations longer than
an hour, about \(80\,\%\) of the CEG have a duration of less than 5
minutes.

>>>>

The distribution of the number of CE events as a function of OIR is
shown in Figure\nobreakspace{}\ref{fig:ovir-distribution}. Apparently,
there is an inverse relation between the frequency of CE events and OIR
with an exponential-like decline. This is expected, as the stronger the
CE events become, the rarer are the conditions favoring the occurrence
of CE events. For the majority (over \(62\%\)) of the CE events the OIR
is below the long term average of \(39.9\,\text{W}/\text{m}^2\), while
about \(8.1\%\) of the events correspond to OIR larger than
\(100\,\text{W}/\text{m}^2\) and up to highest value of
\(412.4\,\text{W}/\text{m}^2\). This distribution is indicative of the
magnitude and the probability of the expected CE events over
Thessaloniki. Similar distribution of CE events, albeit with larger OIR
values, has been reported by \citet{Vamvakas2020}, fot the city of
Patras.This site is located \(2.5^\circ\) closer to the equator. south
of Thessaloniki and is exposed to air masses coming mainly from the
eastern Mediterranean resulting in different cloud patterns, that may
affect the characteristics and magnitude of the CE events.

@@

\caption{Relative frequency distribution of CE events as a function of OIR.}\label{fig:ovir-distribution}
\end{figure}

\hypertarget{duration-of-cloud-enhancement-events}{%
\subsection{Duration of cloud enhancement
events}\label{duration-of-cloud-enhancement-events}}

The duration of the CE events is variable and can last for several
minutes or even more than an hour. In order to study the characteristics
of these consecutive events, we grouped them into bins of increasing
duration in steps of one minute.

We have identified 28468 groups of CE in the whole period of study,
where the group with the longest duration lasted 140 minutes on 07~July
2013. By examining the frequency distribution of the CEG durations
(Figure\nobreakspace{}\ref{fig:ceg-duration-distribution}), we can
conclude, that although we detected some CEG with durations longer than
an hour, about \(80\,\%\) of the CEG have a duration of less than 5
minutes.
